---
title: "Assignment4"
Name : "Maddukuri Janakisrija"
Date : "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
setwd("~/Documents/GitHub/64018_jmadduku/Assignment4")
```

# Transportation Problem:

```{r}
library(lpSolveAPI)
transportation.lp <- read.lp ("transportation.lp")
```

# Printing the Transportation Model
```{r}
transportation.lp
solve(transportation.lp)
```

#To print the optimum solution for Model

```{r}
get.objective(transportation.lp)
```
# The optimum solution for the model is 132,790 which is the combined cost of both production and shipping.


# To print the optimum values from the units produced and shipped from the plant to warehouse.

```{r}
decision_Vars <- get.variables(transportation.lp)
print(paste("optimum values of decision variables are:", as.data.frame(decision_Vars)))
```

# To print the optimum values for the constraints which have feasible solution.

```{r}
const <- get.constraints(transportation.lp)
print(paste("The optimum values of the constraints are:", as.data.frame(const)))
```