---
title : "QMM_Assignment5"
author : "Maddukuri Janakisrija"
date : "11/08/2021"
output : 
pdf_document : default
word_document : default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lpSolveAPI)  # loading package- lpSolveAPI
library(Benchmarking)  # loading package- Benchmarking
library(cowplot)
```

```{r}
dmu1 <- read.lp("dmu1.lp")
dmu1
solve(dmu1)
get.objective(dmu1)
get.variables(dmu1)
```

# The efficiency of the dmu1 is 1 which is maximum

# The outputs of the lp are 5.17241 and 1.12069
# The inputs of the lp are 7.14285 and 0.00 for maximum efficiency.

```{r}
dmu2 <- read.lp ("dmu2.lp")
dmu2
solve(dmu2)
get.objective(dmu2)
get.variables(dmu2)
```

# The efficiency of the dmu2 is 1 which is maximum

# The outputs of the lp are 1.29969 and 6.85889
# The inputs of the lp are 0.00 and 4.76190 for maximum efficiency.

```{r}
dmu3 <- read.lp("dmu3.lp")
dmu3
solve(dmu3)
get.objective(dmu3)
get.variables(dmu3)
```

# The efficiency of the dmu3 is 1 which is maximum

# The outputs of the lp are 1.72413 and 3.73563
# The inputs of the lp are 2.38095 and 0.000 for maximum efficiency.

```{r}
dmu4 <- read.lp("dmu4.lp")
dmu4
solve(dmu4)
get.objective(dmu4)
get.variables(dmu4)
```

# The efficiency of the dmu4 is 1 which is maximum

# The outputs of the lp are 1.92307 and 0.000
# The inputs of the lp are 1.05565 and 1.63817 for maximum efficiency.

```{r}
dmu5 <- read.lp("dmu5.lp")
dmu5
solve(dmu5)
get.objective(dmu5)
get.variables(dmu5)
```

# The efficiency of the dmu5 is 1 which is maximum

# The outputs of the lp are 1.46950 and 7.56796
# The inputs of the lp are 1.59021 and 4.26157 for maximum efficiency.

```{r}
dmu6 <- read.lp("dmu6.lp")
dmu6
solve(dmu6)
get.objective(dmu6)
get.variables(dmu6)
```

# The efficiency of the dmu6 is 1 which is maximum
# The outputs of the lp are 1.46950 and 7.56796
# The inputs of the lp are 1.59021 and 4.26157 for maximum efficiency.
# dmu is not efficient even with the greatest weights.

The inputs and outputs are defined as vector and 2 inputs which are staff hours and supplies and 2 outputs which are Reimbursed Patient_days and Privately Paid Patient_Day.

```{r}
a <- matrix(c(150, 400, 320, 520, 350, 320, 0.2, 0.7, 1.2, 2.0, 1.2, 0.7), ncol = 2)
b <- matrix(c(14000,14000,42000,28000,19000,14000,3500,21000,10500,42000,25000,15000),ncol = 2)
colnames(a) <- c("Staff_Hours", "Supplies")
colnames(b) <- c("Reimbursed Patient_Days","Privately Paid Patient_Days")
print(a)
print(b)
Matrix<- cbind(a,b)
row.names(Matrix) = c("Facility1", "Facility2", "Facility3", "Facility4", "Facility5", "Facility6")
Matrix
```

# 1) Formulate and perform DEA analysis under all DEA assumptions of FDH, CRS, VRS, IRS, DRS, and FRH. 

# 2) Determine the Peers and Lambdas under each of the above assumptions

```{r}
#Now, we are going to formulate and compute the DEA  analysis using FDH.

 #The free disposability assumption stipulates that we can freely discard unnecessary inputs and unwanted outputs

FDH <- dea(a,b, RTS = "fdh")
FDH
peers(FDH)
FDH_Weights <- lambda(FDH)

```
The value of the peer for each facility is always the same as the peer.

 
# Now, we are going to formulate and compute the DEA  analysis using CRS

#Constant returns to scale (CRS) is one of the scaling assumptions. This assumption stipulates that if any possible production combination can arbitrarily be scaled up or down.

```{r}
CRS <- dea(a,b, RTS = "crs")
CRS
#Identify Peers
peers(CRS)
#Identify lambda
CRS_Weights <- lambda(CRS)

```
The results show DMU has 1 which is maximum efficiency for 1,2,3,4  and DMU 5 is 0.9775, DMU 6 0.8675
The peer for 5 and 6 are 1,2,4

# Now, we are going to formulate and compute the DEA  analysis using VRS.

# VRS is one of the scaling assumptions. This Assumptions represents that variable return to scale


```{r}
VRS <- dea(a,b, RTS = "vrs")
VRS
peers(VRS)
VRS_Weights <- lambda(VRS)
```
All facilities 1,2,3,4,5 are efficient except DMU6 which is 0.8963



#Now, we are going to formulate and compute the DEA  analysis using IRS.

#IRS is one of the scaling assumptions It represents that for possible production process we can arbitrarily increase the scale of the operation.

```{r}
IRS <- dea(a,b, RTS = "irs")
IRS
peers(IRS)
IRS_Weights <- lambda(IRS)

```

```{r}

# Now, we are going to formulate and compute the DEA  analysis using DRS.

# DRS is one of the scaling assumptions It represents that for possible production process we can arbitrarily decrease the scale of the operation.

DRS <- dea(a,b, RTS = "drs")
DRS
peers(DRS)
DRS_Weights <- lambda(DRS)

```

```{r}
# Now, we are going to formulate and compute the DEA  analysis using add.

# FRH is a free disposability and replicability hull assumption. It indicates the Additivity (scaling up and down, but only with integers) and free disposability.

FRH <- dea(a,b, RTS="add")
FRH
peers(FRH)
FRH_Weights <- lambda(FRH)

```


```{r}

as.data.frame(Matrix)
DataFrame<- data.frame(CRS = c(1.0000, 1.0000, 1.0000, 1.0000, 0.9775, 0.8675), FDH = c(1, 1, 1 ,1 ,1 ,1), VRS =c(1.0000, 1.0000, 1.0000, 1.0000 ,1.0000, 0.8963), IRS = c(1.0000, 1.0000, 1.0000, 1.0000, 1.0000, 0.8963), DRS = c(1.0000, 1.0000, 1.0000 ,1.0000 ,0.9775, 0.8675), FRH = c(1, 1, 1 ,1, 1, 1))
DataFrame

```

From the above output the Facilities 1,2,3,4 are fully efficient for all the assumptions and Facilities 5,6 are not efficient.

Facility 5 is fully efficient for FDH, VRS, IRS and FRH assumptions.
It is observed that 97.75% efficient for CRS and DRS assumptions.
Facility 6 is fully efficient for FDH and FRS assumptions.
For Facility 6 CRS and DRS assumptions 86.75% efficient.
For Facility 6 IRS and VRS assumptions 89.63% efficient.

# Question 2 : GOAL PRORAMMING

Maximize Z = P - 6C - 3D, where
P = total (discounted) profit over the life of the new products,
C = change (in either direction) in the current level of employment,
D = decrease (if any) in next year’s earnings from the current year’s level.

Profit P is expressed as:

P = 20x1 +15x2 + 25x3

Employment level is expressed as :

6x1 + 4x2 + 5x3 = 50

Next year Earnings goal is expressed as:

8x1 + 7x2 + 5x3 >=75

1) Model_Formulation:

Let us consider
y1 - Employment Level minus the target and
y2 - Next Year Earnings minus the Target
y1+ - Penalty for employment level goal exceeding 50
y1- - Penalty for employment level goal decreasing below 50
y2+ - Exceed the next year earnings
y2- - Penalty for not reaching the next year earnings

y1 = 6x1 + 4x2 + 5x3 - 50
y2 = 8x1 + 7x2 + 5x3 - 75

For Employment level goal
y1 = y1+ - y1- where y1 +, y1 - >=0
y1+ - y1-= 6x1 + 4x2 + 5x3 - 50

For Next year earnings goal
y2 = y2+ - y2- where y2+, y2- >=0
y2+ - y2- = 8x1 + 7x2 + 5x3 -75

Final Formulation is expressed as
Max P = 20x1 +15x2 + 25x3
6x1 + 4x2 + 5x3 - (y1+ - y1-) = 50
8x1 + 7x2 + 5x3 - (y2+ - y2-) = 75

xj >= 0, where j = 1,2,3
yi + >= 0, where i = 1,2
yi - >= 0, where i = 1,2

#2)Managements objective function Objective Function

Maximize Z = P - 6C - 3D
Objective function in terms of x1, x2, x3, y1+, y1- , y2+ and y2-
Max Z = 20x1 +15x2 + 25x3 - 6y1+ - 6y1- - 3y2-
S.T.:
6x1 + 4x2 + 5x3 - y1+ + y1- = 50
8x1 + 7x2 + 5x3 - y2+ + y2- = 75
xj >=0 where j=1,2,3
yi + >=0 where i= 1,2
yi - >=0 where i= 1,2

## 3) Formulate and solve the linear programming model. What are your findings? 

```{r}
library(lpSolveAPI)
Maximum<- read.lp("max.lp")
solve(Maximum)
#Print the model
Maximum
```
```{r}
#To identify the Optimal Solution
get.objective(Maximum)
```

The Optimal Value for the model is 225

```{r}
#To Identify the variables
get.variables(Maximum)
```

When the labor and earnings goals are prioritized, the optimal value can be obtained simply by producing products 2 and 3 at rates of 8.33 and 3.33, respectively. While both the workforce and earnings targets are met.

# Problems Found

According to the Non Preemptive goal programming model, the objective function can be maximized by adding 25 additional people, which is a non-feasible approach for management.
In the preemptive goal programming model using the streamlined approach, we have given the Employment and earnings goals a higher priority than the total profit goal, and the solution indicates that the optimal value has been achieved by achieving all of the goals, as opposed to the previous model where the Employment goal was not achieved.

