---
title: "Integer Programming Exercise"
author : "Maddukuri Janakisrija"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#This rmd file contains the code for assignment 6. 
#The purpose of this assignment is to explore the integer programming formulations and solutions. 

```{r}
library(lpSolveAPI)
```

# By definition, integer programming is Mixed integer programming refers to a model in which just part of the variables must have integer values (MIP). Binary integer programming refers to IP issues using just binary variables (BIP).

1. Consider the following activity-on-arc project network, where the 12 arcs (arrows) represent the 12 activities (tasks) that must be performed to complete the project and the network displays the order in which the activities need to be performed. The number next to each arc (arrow) is the time required for the corresponding activity. Consider the problem of finding the longest path (the largest total time) through this network from start (node 1) to finish (node 9), since the longest path is the critical path.  
 
 
 This assignment has nine nodes. Aij is the time it takes for an action to complete from node i to node j. The earliest start time from node i to node j is Esij. The earliest time to complete an activity is Efij, which is computed as Esij + Aij = Efij. The following are the duration times:

A12 = 5
A13 = 3
A25 = 2
A35 = 3
A58 = 2
A57 = 6
A89 = 7
A96 = 5
A46 = 1
A24 = 4
A47 = 4
A79 = 4

if we assume the largest time to start the next action or activity , then (Esij, Aij, Efij) will occur:

(Es13, A13, Ef13) = (0,3,3)
(Es12, A12, Ef12) = (0,5,5)
(Es25, A25, Ef25) = (5,2,7)
(Es35, A35, Ef35) = (3,3,6)
(Es58, A58, Ef58) = (7,2,9)
(Es89, A89, Ef89) = (9,7,16)
(Es24, A24, Ef24) = (5,4,9)
(Es46, A46, Ef46) = (9,1,10)
(Es69, A69, Ef69) = (10,5,15)
(Es57, A57, Ef57) = (7,6,13)
(Es47, A47, Ef47) = (9,4,13)
(Es79, A79, Ef79) = (13,4,17)

The critical or the longest path is given by the two separate paths of activities:

1 - 2 - 4 - 7 - 9

or

1 - 2 - 5 - 7 - 9

```{r}
# create IP project

cip.ip <- make.lp (nrow = 6, ncol = 7)

# create names for nodes and arcs

arc.names <- c("x12", "x13", "x23", "x35", "x34", "x45", "x56")
node.names <- c("node1", "node2", "node3", "node4", "node5", "node6")

#renaming the IP object

rownames(cip.ip) <- node.names
colnames(cip.ip) <- arc.names

#objective function

time <- c(9, 6, 0, 8, 7, 10, 12)
set.objfn(cip.ip, -1 * time)

# set constraints to the LHS

set.row(cip.ip, 1, c(1, 1), indices = c(1,2))  # node 1 which is starting node
set.row(cip.ip, 2, c(1, -1), indices =  c(1,3)) # node 2 which is an intermediate node
set.row(cip.ip, 3, c(1, 1, -1, -1), indices = c( 2, 3, 4, 5))  # node 3 which is intermediate node
set.row(cip.ip, 4, c(1, -1), indices = c(5, 6))  # node 4 which is intermediate node
set.row(cip.ip, 5, c(1, 1, -1), indices = c(4, 6, 7)) # node 5 which is intermediate node
set.row(cip.ip, 6, 1, indices = 7)  # node 6 is finish node

# set the constraints type
set.constr.type(cip.ip, rep("="), 6)

# set constraints to the RHS
rhs <- c(1, rep(0, 4), 1)
set.rhs(cip.ip, rhs)

# set the type to be binary for all the variables
set.type(cip.ip, 1:7, "binary")

#solve the IP problem
solve(cip.ip)

#get objective value
get.objective(cip.ip)

#get the decision variables
get.variables(cip.ip)

#get the constraint RHS values
get.constraints(cip.ip)

```

each arc has two roles which are the representation as an activity and it also defines the precedence relationships among the activities. The arc 2 to 3 usually represents a dummy activity.

Running the LP formulation on any LP solver, the critical path is:

from node 1 to 2
from node 2 to 4
from node 4 to 7
from node 7 to 9

```{r}

cpm.ip <- make.lp(nrow = 9, ncol = 12) 
#
# nrow is the number of nodes, ncol is the numnber of arcs
#
# We need to create the names for nodes and arcs.
# We then create the arc names first.
arc.names <- c("x12", "x13", "x35", "x58", "x57", "x25", "x24","x47", "x46", "x79", "x69", "x89")
# We next create the node names second.
node.names <- c("node_1", "node_2", "node_3", "node_4", "node_5", "node_6", "node_7", "node_8", "node_9")
# rename the IP object
rownames(cpm.ip) <- node.names
colnames(cpm.ip) <- arc.names
# Using classical LP formulation, we will create the objective function.
##obj function.
#
time <- c(5, 3, 3, 2, 6, 2, 4, 4, 1, 4, 5, 7)
set.objfn(cpm.ip, -1*time)
# The default is min, we must change it to max. If we multiple -1 with the times, it becomes a max problem being solved as a min.
# lp.control(cpm.ip,sense="max")
#
# set constraints LHS
set.row(cpm.ip, 1, c(1, 1), indices = c(1, 2) )     
# node 1 (starting node), x12 + x13 = 1
set.row(cpm.ip, 2, c(1, -1, -1), indices = c(1, 6, 7))    
# node 2 (intermediate node), x12 - x25 - x24 = 0
set.row(cpm.ip, 3, c(1, -1), indices = c(2, 3))  
# node 3 (intermediate node), x13 - x35 = 0
set.row(cpm.ip, 4, c(1, -1, -1), indices = c(7, 8, 9))     
# node 4 (intermediate node), x24 - x47 - x46 = 0
set.row(cpm.ip, 5, c(1, 1, -1, -1), indices = c(3, 6, 4, 5))         
# node 5 (intermediate node), x35 + x25 - x58 - x57 = 0
set.row(cpm.ip, 6, c(1, -1), indices = c(9, 11))         
# node 6 (intermediate node), x46 - x69 = 0
set.row(cpm.ip, 7, c(1, 1, -1), indices = c(5, 8, 10))         
# node 7 (intermediate node), x57 + x47 - x79 = 0
set.row(cpm.ip, 8, c(1, -1), indices = c(4, 12))         
# node 8 (intermediate node), x58 - x89 = 0
set.row(cpm.ip, 9, c(1, 1, 1), indices = c(12, 10, 11))         
# node 9 (finish node), x89 + x79 + x69 = 1
## set constraints type
set.constr.type(cpm.ip, rep("="), 9)
## set constraint RHS
rhs <- c(1, rep(0, 7), 1)
set.rhs(cpm.ip, rhs)
## set all variables type to be binary
set.type(cpm.ip, 1:12, "binary")
## solve the IP problem
solve(cpm.ip)
get.objective(cpm.ip)        
# get objective value
get.variables(cpm.ip)        
# get values of decision variables
get.constraints(cpm.ip)      
# get constraint RHS values
```







